---
title: "Applied Mathematics and Statistics"
subtitle: "Talk 01"
author: "Xiangqi Yi (yixq@jmu.edu.cn)"
date: "2025-02-20"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Why do you choose this course?

## About this course
- General assumptions(limitations) of basic statistical methods
  - Student's t test
  - ANOVA
  - Linear regression
  - Pearson's correlation analysis
  ...
- Beyond linear regression
  - Linear mixed-effects model (LMM)
  - Generalized linear model (GLM)
  - Generalized additive model(GAM)
  ...

## About this course
- Tools
  - R, Rstudio and R markdown
  - Git and github
  -...

##

"Young man, in mathematics you don’t understand things. You just get used to them."

by JOHN VON NEUMANN, to FELIX SMITH

## How to achieve good scores in this course？
### Homework(50%) + Final project (50%)
#### Homework
- Deadline: Wednesday
- `.Rmd` files + others

#### Final project
- Deadline: June 21, 2025 (Week 18)
- `.Rmd` files + others

## About final project
- Topic selection
  - Select one from the topic collection
  - Self-selected topic confirmed by me
- Scoring criteria
  - A complete story
  - Appropriate statistical methods
  - easy or hard
  ...
- ***No tolerance for plagiarism***, but large language models, such as DeepSeek, ChatGPT and Qwen, are acceptable.

## Resources

- [R for Data Science(R4DS)](https://r4ds.hadley.nz): This popular book by Hadley Wickham and Garrett Grolemund covers data manipulation, visualization, and modeling in `R` and `tidyverse`.
- *Mixed Effects Models And Extensions in Ecology with R* 
- [Swirl](https://swirlstats.com): An `R` package that turns your `R` console into an interactive learning environment. 

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Underlying assumptions of linear regression
- Linearity: the independent variables is linearly related to dependent variable.
- Independence: the observations are independent of each other.
- Homoscedasticity: the variance of the residuals (errors) is constant across all levels of the independent variables.
- Normality: the residuals follow a normal distribution.
- No multicollinearity: independent variables are not highly correlated with each other.

# A case of linear regression model
## Import packages
```{r, echo = T, message=FALSE}
# chooseCRANmirror()
# install.packages("tidyverse")
library(tidyverse)
```

## Read data to `dataframe`
```{r, echo=T, message=FALSE}
Clams <- read_table("Data/ZuurDataMixedModelling/Clams.txt")
```
## A glimpse of the `dataframe`
```{r, echo=TRUE}
head(Clams, n = 3) # try glimpse(), summary(), tail(), and slice()
```

## Data transformation and pipe
```{r}
Clams %>%
  mutate(log_AFD = log(AFD),
         log_LENGTH = log(LENGTH)) %>%
  ggplot() + geom_point(aes(log_LENGTH, log_AFD)) +
    facet_wrap(~MONTH)
```


## Homework 01
- Installing `R` and `Rstudio`
- 
